COG-ED Manuscript

### 2021-12-02 06_Paper/COG-ED/Manuscript.Rmd

  # we will add the package versions before submission
  # we also have the word count add-in by Ben Marwick, which can be installed via
  # devtools::install_github("benmarwick/wordcountaddin", type = "source", dependencies = TRUE)

@@

  # remove trials with RT < 200 ms from the n-back data and exclude trials after incorrect trials
  data_RT <- data_nback[-c(which(data_nback$rt < 0.2)), ]
  data_RT <- data_RT[-c(which(data_RT$correct == 0) + 1), ]

@@

  # create new data frames for the specification curve analysis
  # the data frame names are a combination of the letters from each of these categories, indicating how they will be preprocessed:
  #
  # across which dimensions the tranformation and outlier exclusion will be done:
  #   (AA) across subjects, across n-back levels
  #   (AW) across subjects, within n-back levels
  #   (WW) within subjects, within n-back levels
  #   (WA) within subjects, across n-back levels
  # RT transformations:
  #   (R) raw/none
  #   (L) log
  #   (I) inverse
  #   (S) square-root
  # RT outlier exclusion:
  #   (N) none
  #   (2) 2 SDs from the mean
  #   (3) 3 SDs from the mean
  #   (T) 0-200 ms after stimulus onset
  # Post-error trials:
  #   (E) excluded
  #   (N) not excluded
  #
  # Example: the data frame "AWL3E" has log-transformed RT data with no data points beyond 3 SD
  # from the mean and with no trials that followed misses/false alarms
  #
  # Some combinations are nonsensical, so we will not include them: Raw data sets (__R___) with either no outlier
  # exclusion (___N_) or time-based outlier exclusion (___T_) do not need the different dimensions (A/W), and
  # time-based exclusion is nonsensical for transformed data sets.

@@

          # compute mean and standard deviation

          pipemean <- mean(pipelines_data[[i]]$rt, na.rm = TRUE)
          pipesd <- sd(pipelines_data[[i]]$rt, na.rm = TRUE)

@@

            pipelines_data[[i]] <- pipelines_data[[i]][c(which(pipelines_data[[i]]$rt < (pipemean + 2 * pipesd)
                                                               & pipelines_data[[i]]$rt > (pipemean - 2 * pipesd))), ]

@@

            pipelines_data[[i]] <- pipelines_data[[i]][c(which(pipelines_data[[i]]$rt < (pipemean + 3 * pipesd)
                                                               & pipelines_data[[i]]$rt > (pipemean - 3 * pipesd))), ]
          }

@@

            # compute mean and standard deviation

            pipemean <- mean(pipelines_data[[i]]$rt[pipelines_data[[i]]$level == x], na.rm = TRUE)
            pipesd <- sd(pipelines_data[[i]]$rt[pipelines_data[[i]]$level == x], na.rm = TRUE)

@@

              ifelse(length(which(pipelines_data[[i]]$level == x &
                                    (pipelines_data[[i]]$rt > (pipemean + 2 * pipesd) | pipelines_data[[i]]$rt < (pipemean - 2 * pipesd)))) > 0,
                     pipelines_data[[i]] <-
                       pipelines_data[[i]][-c(which(pipelines_data[[i]]$level == x &
                                                      (pipelines_data[[i]]$rt > (pipemean + 2 * pipesd) | pipelines_data[[i]]$rt < (pipemean - 2 * pipesd)))), ],
                     pipelines_data[[i]] <- pipelines_data[[i]])

@@

              ifelse(length(which(pipelines_data[[i]]$level == x &
                                    (pipelines_data[[i]]$rt > (pipemean + 3 * pipesd) | pipelines_data[[i]]$rt < (pipemean - 3 * pipesd)))) > 0,
                     pipelines_data[[i]] <-
                       pipelines_data[[i]][-c(which(pipelines_data[[i]]$level == x &
                                                    (pipelines_data[[i]]$rt > (pipemean + 3 * pipesd) | pipelines_data[[i]]$rt < (pipemean - 3 * pipesd)))), ],
                     pipelines_data[[i]] <- pipelines_data[[i]])
            }
        }

@@

            # compute mean and standard deviation

            pipemean <- mean(pipelines_data[[i]]$rt[pipelines_data[[i]]$level == x & pipelines_data[[i]]$subject == pipesubject], na.rm = TRUE)
            pipesd <- sd(pipelines_data[[i]]$rt[pipelines_data[[i]]$level == x & pipelines_data[[i]]$subject == pipesubject], na.rm = TRUE)

@@

              ifelse(length(which(pipelines_data[[i]]$level == x & pipelines_data[[i]]$subject == pipesubject &
                                    (pipelines_data[[i]]$rt > (pipemean + 2 * pipesd) | pipelines_data[[i]]$rt < (pipemean - 2 * pipesd)))) > 0,
                     pipelines_data[[i]] <-
                       pipelines_data[[i]][-c(which(pipelines_data[[i]]$level == x & pipelines_data[[i]]$subject == pipesubject &
                                                      (pipelines_data[[i]]$rt > (pipemean + 2 * pipesd) | pipelines_data[[i]]$rt < (pipemean - 2 * pipesd)))), ],
                     pipelines_data[[i]] <- pipelines_data[[i]])

@@

              ifelse(length(which(pipelines_data[[i]]$level == x & pipelines_data[[i]]$subject == pipesubject &
                                    (pipelines_data[[i]]$rt > (pipemean + 3 * pipesd) | pipelines_data[[i]]$rt < (pipemean - 3 * pipesd)))) > 0,
                     pipelines_data[[i]] <-
                       pipelines_data[[i]][-c(which(pipelines_data[[i]]$level == x & pipelines_data[[i]]$subject == pipesubject &
                                                      (pipelines_data[[i]]$rt > (pipemean + 3 * pipesd) | pipelines_data[[i]]$rt < (pipemean - 3 * pipesd)))), ],
                     pipelines_data[[i]] <- pipelines_data[[i]])
            }
          }
        }

@@

            # compute mean and standard deviation

            pipemean <- mean(pipelines_data[[i]]$rt[pipelines_data[[i]]$subject == pipesubject], na.rm = TRUE)
            pipesd <- sd(pipelines_data[[i]]$rt[pipelines_data[[i]]$subject == pipesubject], na.rm = TRUE)

@@

              ifelse(length(which(pipelines_data[[i]]$subject == pipesubject &
                                    (pipelines_data[[i]]$rt > (pipemean + 2 * pipesd) | pipelines_data[[i]]$rt < (pipemean - 2 * pipesd)))) > 0,
                     pipelines_data[[i]] <-
                       pipelines_data[[i]][-c(which(pipelines_data[[i]]$subject == pipesubject &
                                                      (pipelines_data[[i]]$rt > (pipemean + 2 * pipesd) | pipelines_data[[i]]$rt < (pipemean - 2 * pipesd)))), ],
                     pipelines_data[[i]] <- pipelines_data[[i]])

@@

              ifelse(length(which(pipelines_data[[i]]$subject == pipesubject &
                                    (pipelines_data[[i]]$rt > (pipemean + 3 * pipesd) | pipelines_data[[i]]$rt < (pipemean - 3 * pipesd)))) > 0,
                     pipelines_data[[i]] <-
                       pipelines_data[[i]][-c(which(pipelines_data[[i]]$subject == pipesubject &
                                                      (pipelines_data[[i]]$rt > (pipemean + 3 * pipesd) | pipelines_data[[i]]$rt < (pipemean - 3 * pipesd)))), ],
                     pipelines_data[[i]] <- pipelines_data[[i]])
            }
        }

@@

    # the 200-ms-exclusion and the post-error-exclusion are done outside of the A/A-A/W-W/W-W/A-conditions because they are independent of dimension:

    # exclude RTs fast than 200 ms after stimulus onset if the fourth letter in the data frame name is a "T"

@@

      pipelines_data[[i]] <- pipelines_data[[i]][-c(which(pipelines_data[[i]]$rt < 0.2)), ]

@@

We aim to conduct all analysis as described in Westbrook et al. [-@Westbrook2013], but the level of detail was not always sufficient, so there might be deviations regarding data cleaning and degrees of freedom. The performance measure *d'* will be computed as the difference of the *z*-transformed hit rate and the *z*-transformed false alarm rate [@Macmillan1990]. Reaction time (RT) data will be trimmed by excluding all trials with responses faster than 200 ms, as the relevant cognitive processes cannot have been completed before [@Whelan2008; @Berger2021]. Error- and post-error trials will be excluded, because RT on the latter is longer due to more cautious behavior [@Dutilh2012; @Houtman2012].
Declining performance will be investigated by calculating a repeated measures analysis of variance (rmANOVA) with three paired contrasts comparing *d'* between two levels of 2-, 3-, and 4-back at a time.
Another rmANOVA with three paired contrasts will be computed to compare the mean RT between two levels of 2-, 3-, and 4-back at a time.
To investigate changes in NASA-TLX ratings, six rmANOVAs will be computed, one for each NASA-TLX subscale, and each with six paired contrasts comparing the ratings between two levels of 1-, 2-, 3-, and 4-back at a time.
For each ED round, SVs will be calculated by adding or subtracting 0.015625 from the last monetary value of the flexible level, depending on the participant's last choice.
Then, these final monetary values will be divided by 2€, and the SV of each level per participant will be computed by averaging all final values of each level, regardless of whether it was fixed or flexible.
A rmANOVA with six paired contrasts will be computed, comparing the SVs between two levels of 1-, 2-, 3-, and 4-back at a time.
Tukey method will be used for the paired contrasts of each rmANOVA, including *p*-value adjustment.

To determine the influence of task performance in the association of SVs and n-back level, we will set up a multi-level model.
**MLM explanation goes here** To ensure the validity of the multi-level model, we will conduct a specification curve analysis [@Simonsohn2020], which will include 92 possible preprocessing pipelines of the RT data.
These pipelines specify which transformation was applied (none, log, inverse, or square-root), which outliers were excluded (none, $2 SD$ from the mean, $3 SD$ from the mean, RTs below 200 ms), whether post-error trials were excluded or not, and across which dimensions the transformations and exclusions were applied (across/within subjects and across/within n-back levels).
The multi-level model will be run with each of the 92 pipelines, which will also include our main pipeline (untransformed data, exclusion of RTs below 200 ms, removal of post-error trials).
The ratio of pipelines that lead to significant versus non-significant effects will provide an indication of how robust the effect actually is.

>>>>

  # we will add the package versions before submission

@@

  # remove trials with RT < 100 ms from the n-back data and exclude trials after incorrect trials
  data_RT <- data_nback[-c(which(data_nback$rt < 0.1)), ]
  data_RT <- data_RT[-c(which(data_RT$correct == 0) + 1), ]

@@

  # create new data frames for the specification curve analysis
  # the data frame names are a combination of the letters from each of these categories, indicating how they will be preprocessed:
  #
  # across which dimensions the tranformation and outlier exclusion will be done:
  #   (AA) across subjects, across n-back levels
  #   (AW) across subjects, within n-back levels
  #   (WW) within subjects, within n-back levels
  #   (WA) within subjects, across n-back levels
  # RT transformations:
  #   (R) raw/none
  #   (L) log
  #   (I) inverse
  #   (S) square-root
  # RT outlier exclusion:
  #   (N) none
  #   (2) 2 SDs from the mean
  #   (3) 3 SDs from the mean
  #   (T) 0-100 ms after stimulus onset
  # Post-error trials:
  #   (E) excluded
  #   (N) not excluded
  #
  # Example: the data frame "AWL3E" has log-transformed RT data with no data points beyond 3 SD
  # from the mean and with no trials that followed misses/false alarms
  #
  # Some combinations are nonsensical, so we will not include them: Raw data sets (__R___) with either no outlier
  # exclusion (___N_) or time-based outlier exclusion (___T_) do not need the different dimensions (A/W), and
  # time-based exclusion is nonsensical for transformed data sets.

@@

          # compute median and median absolute deviation

          pipemedian <- median(pipelines_data[[i]]$rt, na.rm = TRUE)
          pipemad <- mad(pipelines_data[[i]]$rt, na.rm = TRUE)

@@

            pipelines_data[[i]] <- pipelines_data[[i]][c(which(pipelines_data[[i]]$rt < (pipemedian + 2 * pipemad)
                                                               & pipelines_data[[i]]$rt > (pipemedian - 2 * pipemad))), ]

@@

            pipelines_data[[i]] <- pipelines_data[[i]][c(which(pipelines_data[[i]]$rt < (pipemedian + 3 * pipemad)
                                                               & pipelines_data[[i]]$rt > (pipemedian - 3 * pipemad))), ]
          }

@@

            # compute median and median absolute deviation

            pipemedian <- median(pipelines_data[[i]]$rt[pipelines_data[[i]]$level == x], na.rm = TRUE)
            pipemad <- mad(pipelines_data[[i]]$rt[pipelines_data[[i]]$level == x], na.rm = TRUE)

@@

              ifelse(length(which(pipelines_data[[i]]$level == x &
                                    (pipelines_data[[i]]$rt > (pipemedian + 2 * pipemad) | pipelines_data[[i]]$rt < (pipemedian - 2 * pipemad)))) > 0,
                     pipelines_data[[i]] <-
                       pipelines_data[[i]][-c(which(pipelines_data[[i]]$level == x &
                                                      (pipelines_data[[i]]$rt > (pipemedian + 2 * pipemad) | pipelines_data[[i]]$rt < (pipemedian - 2 * pipemad)))), ],
                     pipelines_data[[i]] <- pipelines_data[[i]])

@@

              ifelse(length(which(pipelines_data[[i]]$level == x &
                                    (pipelines_data[[i]]$rt > (pipemedian + 3 * pipemad) | pipelines_data[[i]]$rt < (pipemedian - 3 * pipemad)))) > 0,
                     pipelines_data[[i]] <-
                       pipelines_data[[i]][-c(which(pipelines_data[[i]]$level == x &
                                                    (pipelines_data[[i]]$rt > (pipemedian + 3 * pipemad) | pipelines_data[[i]]$rt < (pipemedian - 3 * pipemad)))), ],
                     pipelines_data[[i]] <- pipelines_data[[i]])
            }
        }

@@

            # compute median and median absolute deviation

            pipemedian <- median(pipelines_data[[i]]$rt[pipelines_data[[i]]$level == x & pipelines_data[[i]]$subject == pipesubject], na.rm = TRUE)
            pipemad <- mad(pipelines_data[[i]]$rt[pipelines_data[[i]]$level == x & pipelines_data[[i]]$subject == pipesubject], na.rm = TRUE)

@@

              ifelse(length(which(pipelines_data[[i]]$level == x & pipelines_data[[i]]$subject == pipesubject &
                                    (pipelines_data[[i]]$rt > (pipemedian + 2 * pipemad) | pipelines_data[[i]]$rt < (pipemedian - 2 * pipemad)))) > 0,
                     pipelines_data[[i]] <-
                       pipelines_data[[i]][-c(which(pipelines_data[[i]]$level == x & pipelines_data[[i]]$subject == pipesubject &
                                                      (pipelines_data[[i]]$rt > (pipemedian + 2 * pipemad) | pipelines_data[[i]]$rt < (pipemedian - 2 * pipemad)))), ],
                     pipelines_data[[i]] <- pipelines_data[[i]])

@@

              ifelse(length(which(pipelines_data[[i]]$level == x & pipelines_data[[i]]$subject == pipesubject &
                                    (pipelines_data[[i]]$rt > (pipemedian + 3 * pipemad) | pipelines_data[[i]]$rt < (pipemedian - 3 * pipemad)))) > 0,
                     pipelines_data[[i]] <-
                       pipelines_data[[i]][-c(which(pipelines_data[[i]]$level == x & pipelines_data[[i]]$subject == pipesubject &
                                                      (pipelines_data[[i]]$rt > (pipemedian + 3 * pipemad) | pipelines_data[[i]]$rt < (pipemedian - 3 * pipemad)))), ],
                     pipelines_data[[i]] <- pipelines_data[[i]])
            }
          }
        }

@@

            # compute median and median absolute deviation

            pipemedian <- median(pipelines_data[[i]]$rt[pipelines_data[[i]]$subject == pipesubject], na.rm = TRUE)
            pipemad <- mad(pipelines_data[[i]]$rt[pipelines_data[[i]]$subject == pipesubject], na.rm = TRUE)

@@

              ifelse(length(which(pipelines_data[[i]]$subject == pipesubject &
                                    (pipelines_data[[i]]$rt > (pipemedian + 2 * pipemad) | pipelines_data[[i]]$rt < (pipemedian - 2 * pipemad)))) > 0,
                     pipelines_data[[i]] <-
                       pipelines_data[[i]][-c(which(pipelines_data[[i]]$subject == pipesubject &
                                                      (pipelines_data[[i]]$rt > (pipemedian + 2 * pipemad) | pipelines_data[[i]]$rt < (pipemedian - 2 * pipemad)))), ],
                     pipelines_data[[i]] <- pipelines_data[[i]])

@@

              ifelse(length(which(pipelines_data[[i]]$subject == pipesubject &
                                    (pipelines_data[[i]]$rt > (pipemedian + 3 * pipemad) | pipelines_data[[i]]$rt < (pipemedian - 3 * pipemad)))) > 0,
                     pipelines_data[[i]] <-
                       pipelines_data[[i]][-c(which(pipelines_data[[i]]$subject == pipesubject &
                                                      (pipelines_data[[i]]$rt > (pipemedian + 3 * pipemad) | pipelines_data[[i]]$rt < (pipemedian - 3 * pipemad)))), ],
                     pipelines_data[[i]] <- pipelines_data[[i]])
            }
        }

@@

    # the 100-ms-exclusion and the post-error-exclusion are done outside of the A/A-A/W-W/W-W/A-conditions because they are independent of dimension:

    # exclude RTs fast than 100 ms after stimulus onset if the fourth letter in the data frame name is a "T"

@@

      pipelines_data[[i]] <- pipelines_data[[i]][-c(which(pipelines_data[[i]]$rt < 0.1)), ]

@@

We aim to conduct all analysis as described in Westbrook et al. [-@Westbrook2013], but the level of detail was not always sufficient, so there might be deviations regarding data cleaning and degrees of freedom. The performance measure *d'* will be computed as the difference of the *z*-transformed hit rate and the *z*-transformed false alarm rate [@Macmillan1990]. Reaction time (RT) data will be trimmed by excluding all trials with responses faster than 100 ms, as the relevant cognitive processes cannot have been completed before [@Whelan2008; @Berger2021]. Error- and post-error trials will be excluded, because RT on the latter is longer due to more cautious behavior [@Dutilh2012; @Houtman2012].
Declining performance will be investigated by calculating a repeated measures analysis of variance (rmANOVA) with three paired contrasts comparing *d'* between two levels of 2-, 3-, and 4-back at a time.
Another rmANOVA with three paired contrasts will be computed to compare the mean RT between two levels of 2-, 3-, and 4-back at a time.
To investigate changes in NASA-TLX ratings, six rmANOVAs will be computed, one for each NASA-TLX subscale, and each with six paired contrasts comparing the ratings between two levels of 1-, 2-, 3-, and 4-back at a time.
For each ED round, SVs will be calculated by adding or subtracting 0.015625 from the last monetary value of the flexible level, depending on the participant's last choice.
Then, these final monetary values will be divided by 2€, and the SV of each level per participant will be computed by averaging all final values of each level, regardless of whether it was fixed or flexible.
A rmANOVA with six paired contrasts will be computed, comparing the SVs between two levels of 1-, 2-, 3-, and 4-back at a time.
Tukey method will be used for the paired contrasts of each rmANOVA, including *p*-value adjustment.

To determine the influence of task performance in the association of SVs and n-back level, we will set up a multi-level model.
**MLM explanation goes here** To ensure the validity of the multi-level model, we will conduct a specification curve analysis [@Simonsohn2020], which will include 92 possible preprocessing pipelines of the RT data.
These pipelines specify which transformation was applied (none, log, inverse, or square-root), which outliers were excluded (none, $2 SD$ from the mean, $3 SD$ from the mean, RTs below 100 ms), whether post-error trials were excluded or not, and across which dimensions the transformations and exclusions were applied (across/within subjects and across/within n-back levels).
The multi-level model will be run with each of the 92 pipelines, which will also include our main pipeline (untransformed data, exclusion of RTs below 100 ms, removal of post-error trials).
The ratio of pipelines that lead to significant versus non-significant effects will provide an indication of how robust the effect actually is.
